---
title: "Project to predict and quantify the manner in which the exercise was done"
author: "Mayank Pundir"
date: '`r Sys.Date()`'
---
##Executive summary:-
This report provides an analysis and evaluation of the data collected to predict which is better exercise out of 5 different ways to quantify how well they do it.
              The method of analysis includes prediction model and exploratory data analyses.The data was extracted from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <http://groupware.les.inf.puc-rio.br/har>.
              
##Data 
The training data for this project are available here: 
<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv>
The test data are available here: 
<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv>

##Exploratory data analyses:-
Lets first check the data content and datatype for the training dataset
```{r,echo=TRUE,message=FALSE,warning=FALSE,cache=TRUE}
temp <- tempfile()
url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
download.file(url,temp,method="curl")
#Replace blank fields to NA
file <- read.csv(temp,header = TRUE,sep = ",",na.strings = c("NA", ""))
dim(file)
```

Before starting with the analysis we will do some data cleaning to the training data.We will be removing fields having sum <> 0 to make data more clean and accurate for prediction.

```{r,echo=TRUE,message=FALSE,warning=FALSE,cache=TRUE}
#Select colums having sum=0 
test <- file[,colSums(is.na(file))==0]
#Remove columns not need for predection like X,user_name,raw_timestamp_part_1,raw_timestamp_part_2,
#cvtd_timestap,new_window and num_window
final <- test[,!names(test) %in% names(test[,c(1,2,3,4,5,6,7)])]

dim(final)
```

Now since the data has been cleaned up and reduced to reqired fields we will do cross data validation by breaking the training data into training and test data.

##cross data validation
```{r,echo=TRUE,message=FALSE,warning=FALSE,cache=TRUE}
library(caret);library(kernlab);
#Breaking data into 75% training and 25% testing data for cross data validation
intain <- createDataPartition(y=final$classe,p=0.75,list=FALSE)
training <- final[intain,]
testing <- final[-intain,]
dim(training);dim(testing)
```

###Create Fit Model
We will create a fit model using training data with method= Knn(K-nearest neighbor)
```{r,echo=TRUE,message=FALSE,warning=FALSE,cache=TRUE}
modelfit <- train(classe ~ .,data=training,method="knn",preProcess = c("center", "scale"),
                  tuneLength = 10,
                  trControl = trainControl(method = "cv"))
modelfit
modelfit$finalModel
prediction <- predict(modelfit,newdata=testing)
```

After creating the fit model we will do prediction on testing data created and create confusionMatrix.
```{r,echo=TRUE,message=FALSE,warning=FALSE,cache=TRUE}
modelfit$finalModel
prediction <- predict(modelfit,newdata=testing)
confusionMatrix(prediction,testing$classe)
```

Below is the Plot created showing all the feature of outcome classe against other fields.

```{r ,echo=TRUE,message=FALSE,warning=FALSE,cache=TRUE}
library(knitr)
featurePlot(x=testing[,grep("total",names(testing))],y=testing$classe,plot="pairs")
```

##Out of sample error
As from above you can see the out of sample error Accuracy = 0.9676

##Applying Above prediction to test
```{r ,echo=TRUE,message=FALSE,warning=FALSE,cache=TRUE}
temp <- tempfile()
url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(url,temp,method="curl")
test <- read.csv(temp,header = TRUE,sep = ",",na.strings = c("NA", ""))
x <- as.character(predict(modelfit,newdata=test))
```

##Results
```{r ,echo=TRUE,message=FALSE,warning=FALSE,cache=TRUE}
print(x)
```